---
title: "braidHymo package"
output:
  rmarkdown::html_vignette:
    df_print: kable
bibliography: braidHymoBib.bib
vignette: >
  %\VignetteIndexEntry{braidHymo package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(braidHymo)
library(dplyr)
```

# Data format

**One type** (1 x `rivdata`) or **two types of files** (n x `rivdata` + `rivers`) must be provided: see details below.

## File `rivdata`

There must be as many `rivdata` files as there are rivers considered.

One `rivdata` file corresponds to

- **1 row per sampling point** with altitude measurement
- **multiple columns**, 2 of which are non-optional and should correspond to:
    + a **cross-section ID** for the point (typically named `ID`)
    + an **altitude measure** for the point (typically named `RASTERVALU`)


## Files `rivers` (optional)

To calculate the indices on **multiple rivers**, a table `rivers` containing information about the rivers should be provided.

The table should contain 

- **n rows** (as many as there are `rivdata` filepaths) 
- **multiple columns**, 3 of which are non-optional and should correspond to:
    + the **catchment area** *in km²* (typically named `area`)
    + the **space between points on each transect** *in m* (typically named `points_space`)
    + the **path to corresponding `rivdata` filepath** (typically named `filepath`)

Other columns might refer to e.g. river name, year, etc.

```{r river_info}
rivers=readr::read_csv("../data-raw/rivers.csv") %>%
    mutate(filepath=paste0("../data-raw/",filepath)) %>% 
    mutate(year=as.factor(year))

# Display the resulting table:
rivers
```

# One river: apply braidHymo_one()

## First, read and clean `rivdata` with braidHymo_read()

```{r read_drac}
data_Drac=braidHymo_read("../data-raw/Drac_Chabottes_2018.txt") 

# Display the first lines of the resulting table:
head(data_Drac)
```

## Then, calculate indices for this river with braidHymo_one():

```{r calc_drac}
result_Drac=braidHymo_one(data_Drac,
                          area=253,
                          points_space=1)

# Display the first lines of the resulting table:
head(result_Drac)
```

# Multiple rivers: apply braidHymo()

Here, we generalise this to 2 rivers.

The function `braidHymo()` retrieves from table `rivers` all the necessary information about rivers, and paths to rivdata filepaths.

```{r braidHymo}
result=braidHymo(rivers)

# Display the first lines of the resulting table:
head(result)
```


It is then possible to graphically display the results for these rivers using function `braidHymo_plot()`:

```{r plot_BRIstar,fig.width=8}
braidHymo_plot(result,index="BRI*", position=year, color=river)
```


```{r plot_Wstar,fig.width=8}
braidHymo_plot(result,index="W*", position=year, color=river)
```

---
nocite: '@*'
---


# References


**When using this package, please refer to @lise_devreux_2021_5796184.**
